\begin{itemize}
    \item [\cite{calleja2022structured}] \textbf{Calleja-Solanas, V.}, Khalil, N., Gómez-Gardeñes, J., Her- nández-García, E., and Meloni, S. Structured interactions as a stabilizing mechanism for competitive ecological communities. Physical Review E, 106(6), 064307 (2022).
    
    \item [\cite{calleja2023predictors}] \textbf{Calleja-Solanas, V.}, Hernández-García, E., and Meloni, S. Structural predictors of species survival in complex ecological communities  \textit{Submitted} (2023).

    \item [\cite{calleja2021quantifying}] \textbf{Calleja-Solanas, V.}, Pigani, E., Palazzi, M. J., Solé-Ribalta, A., Suweis, S., Borge-Holthoefer, J., and Meloni, S. Quantifying the drivers behind collective attention in information ecosystems. Journal of Physics: Complexity, 2(4), 045014 (2021).

    \item [\cite{calleja2023macropatterns}]  \textbf{Calleja-Solanas, V.}, Palazzi, M. J., Plata, C. A., Solé-Ribalta, A., Borge-Holthoefer, J.,  Suweis, S. and Meloni, S. Universal macropatterns in information ecosystems \textit{Submitted} (2023).
\end{itemize}

Other results related to the previous work and also developed during the Ph.D. are:
\begin{itemize}
    \item [\cite{palazzi2021ecological}] Palazzi, M. J., Solé-Ribalta, A., \textbf{Calleja-Solanas, V.}, Meloni, S., Plata, C. A., Suweis, S., and Borge-Holthoefer, J.  An ecological approach to structural flexibility in online communication systems. Nature communications, 12(1), 1-11 (2021).
    
    \item [\cite{plata2021neutral}] Plata, C. A., Pigani, E., Azaele, S., \textbf{Calleja-Solanas, V.}, Palazzi, M. J., Solé-Ribalta, A., Borge-Holthoefer, J., Meloni, S. and Suweis, S.  Neutral theory for competing attention in social networks. Physical Review Research, 3(1), 013070 (2021).
\end{itemize}


